\pkg{webchem}: An \proglang{R} Package to Retrieve Chemical Information from the Web

### 2020-01-30 revision3_2581/paper.tex
removed and replaced Physprop partition coef data with PubChem data

\begin{description}
\item[NIH Chemical Identifier Resolver (CIR)]{\citep{cir} A web
    service that converts from and to various chemical identifiers. The database holds millions of identifier combinations.}
\item[Chemical Translation Service (CTS)]{
  \citep{Wohlgemuth_Haldiya_Willighagen_Kind_Fiehn_2010} A web
    service that converts from and to various chemical identifiers.}
\item[ETOX]{\citep{etox} Information System Ecotoxicology and
    Environmental Quality Targets by the German Federal Environmental
    Agency. Provides basic identifiers, synonyms, ecotoxicological
    data for over 64,000 entries and quality targets for different countries.}
\item[PAN Pesticide Database]{ \citep{pan} Information on pesticides --
    provides basic identifiers, ecotoxicological data, chemical
    properties, uses and regulatory status for 6,500 pesticides.}
\item[SRC Physprop]{\citep{physprop} Contains physical properties for
    over 41,000 chemicals.  Physical properties collected from a wide
    variety of sources including experimental and modeled values.}
\item[PubChem]{\citep{Kim_2016} PubChem is a public repository for
    information on more than 250 million chemical substances, providing identifiers,
    properties and synonyms. We use an interface to the PUG-REST web
    service \citep{Kim_Thiessen_Bolton_Bryant_2015}.}
\item[Wikidata]{\citep{wiki} Wikipedia contains information for over
    210,000 chemicals
    \citep{Ertl_Patiny_Sander_Rufener_Zasso_2015}. Currently
    \pkg{webchem} can only query chemical identifiers.}
\item[Compendium of Pesticide Common Names]{\citep{wood} The
    compendium provides information on pesticide common names,
    identifiers and classification. The compendium contains more than 1800 active ingredients and more than 350 ester and salt derivatives. }
\item[ChemID\emph{plus}]{\citep{Tomasulo_2002} is a large web-based
    database provided by the National Library of Medicine. It contains
    identifiers, synonyms, toxicological data and chemical
    properties for over 420,000 records.}
\item[ChemSpider]{\citep{pence_chemspider:_2010} is a free chemical
    structure database providing access to over 67 million
    structures from hundreds of data sources. It provides identifiers, properties and can also be
    used to convert identifiers.}
\item[OPSIN]{\citep{Lowe_Corbett_Murray-Rust_Glen_2011} The Open
    Parser for Systematic IUPAC nomenclature is a chemical name
    interpreter and provides InChI and SMILES identifiers.}
\item[Flavornet]{\citep{flavornet} Flavornet is a compilation of 738 aroma compounds found in human odor space.}
\item[NIST]{\citep{nist} The NIST Chemistry WebBook provides access to data compiled and distributed by NIST under the Standard Reference Data Program. The WebBook provides chemical and physical property data on over 40,000 compounds.}
\item[ChEBI]{\citep{chebi} Chemical Entities of Biological Interest (ChEBI) is a freely available dictionary of molecular entities focused on "small" chemical compounds. The database contains more than 56,000 compounds.}

@@

\subsection[Querying partitioning coefficients]{Querying partitioning coefficients}
Some data sources also provide data on chemical properties that can be
queried.  Here we query for the \code{lc50} data the
$\mathrm{log}~P_{oct/wat}$ from the SRC Physprop database to build a
simple quantitative structure-activity relationship (QSAR) to predict
toxicity.
%
\begin{CodeChunk}
\begin{CodeInput}
R> pp_data <- pp_query(lc50$cas)
\end{CodeInput}
\end{CodeChunk}
%
The database contains predicted and experimental values.
Extracting $\mathrm{log}~P_{oct/wat}$ from the data object is slightly more complicated,  
because i) for some compounds no data could be found and ii) the data-object has a more complex structure (a data frame within a list).
%
\begin{CodeChunk}
\begin{CodeInput}
R> lc50$logp <- sapply(pp_data, function(y) {
+    if (length(y) == 1 && is.na(y))
+      return(NA)
+    y$prop$value[y$prop$variable == "Log P (octanol-water)"]
+  })
\end{CodeInput}
\end{CodeChunk}
%
We opted for this more complex approach, because the information available is very diverse and we cannot provide an extractor function for each purpose.
Moreover, it provides users with high flexibility regarding organization of their data. 
Nevertheless, in the documentation of each function we provide examples on how to extract more complicated parts of the data.
The resulting data and model are displayed in Figure~\ref{fig:fig3}.

\begin{figure}[t!]
  \centering
  \includegraphics[width=0.7\textwidth]{plot_qsar-1} 
  \caption{Simple QSAR for predicting log LC\textsubscript{50} of
    pesticides by $\mathrm{log}~P$.  $\mathrm{log}~P$ values have been
    retrieved from the SRC Physprop database (97 experimental data, 9
    estimated data and 18 substances without data).  Blue line
    indicates the regression model
    ($\mathrm{log~LC}_{50} = 2.88\ensuremath{-0.37}~\mathrm{log}~P$,
    RMSE = 1.45).}
\label{fig:fig3}
\end{figure}

>>>>

\begin{description}
\item[NIH Chemical Identifier Resolver (CIR)]{\citep{cir} A web
    service that converts from and to various chemical identifiers. The database holds millions of identifier combinations.}
\item[Chemical Translation Service (CTS)]{
  \citep{Wohlgemuth_Haldiya_Willighagen_Kind_Fiehn_2010} A web
    service that converts from and to various chemical identifiers.}
\item[ETOX]{\citep{etox} Information System Ecotoxicology and
    Environmental Quality Targets by the German Federal Environmental
    Agency. Provides basic identifiers, synonyms, ecotoxicological
    data for over 64,000 entries and quality targets for different countries.}
\item[PAN Pesticide Database]{ \citep{pan} Information on pesticides --
    provides basic identifiers, ecotoxicological data, chemical
    properties, uses and regulatory status for 6,500 pesticides.}
\item[PubChem]{\citep{Kim_2016} PubChem is a public repository for
    information on more than 250 million chemical substances, providing identifiers,
    properties and synonyms. We use an interface to the PUG-REST web
    service \citep{Kim_Thiessen_Bolton_Bryant_2015}.}
\item[Wikidata]{\citep{wiki} Wikipedia contains information for over
    210,000 chemicals
    \citep{Ertl_Patiny_Sander_Rufener_Zasso_2015}. Currently
    \pkg{webchem} can only query chemical identifiers.}
\item[Compendium of Pesticide Common Names]{\citep{wood} The
    compendium provides information on pesticide common names,
    identifiers and classification. The compendium contains more than 1800 active ingredients and more than 350 ester and salt derivatives. }
\item[ChemID\emph{plus}]{\citep{Tomasulo_2002} is a large web-based
    database provided by the National Library of Medicine. It contains
    identifiers, synonyms, toxicological data and chemical
    properties for over 420,000 records.}
\item[ChemSpider]{\citep{pence_chemspider:_2010} is a free chemical
    structure database providing access to over 67 million
    structures from hundreds of data sources. It provides identifiers, properties and can also be
    used to convert identifiers.}
\item[OPSIN]{\citep{Lowe_Corbett_Murray-Rust_Glen_2011} The Open
    Parser for Systematic IUPAC nomenclature is a chemical name
    interpreter and provides InChI and SMILES identifiers.}
\item[Flavornet]{\citep{flavornet} Flavornet is a compilation of 738 aroma compounds found in human odor space.}
\item[NIST]{\citep{nist} The NIST Chemistry WebBook provides access to data compiled and distributed by NIST under the Standard Reference Data Program. The WebBook provides chemical and physical property data on over 40,000 compounds.}
\item[ChEBI]{\citep{chebi} Chemical Entities of Biological Interest (ChEBI) is a freely available dictionary of molecular entities focused on "small" chemical compounds. The database contains more than 56,000 compounds.}

@@

\subsection[Querying partitioning coefficients]{Querying partitioning coefficients}
Some data sources also provide data on chemical properties that can be
queried.  Here we query for the \code{lc50} data predicted
$\mathrm{log}~P_{oct/wat}$ from the PubChem database to build a
simple quantitative structure-activity relationship (QSAR) to predict
toxicity. The resulting data and model are displayed in Figure~\ref{fig:fig3}.
%
\begin{CodeChunk}
\begin{CodeInput}
R> cid <- get_cid(lc50$cas)
R> pc_data <- pc_prop(cid)
R> lc50$logp <- pc_data$XLogP
\end{CodeInput}
\end{CodeChunk}
%

\begin{figure}[t!]
  \centering
  \includegraphics[width=0.7\textwidth]{plot_qsar-1} 
  \caption{Simple QSAR for predicting log LC\textsubscript{50} of
    pesticides by $\mathrm{log}~P$.  $\mathrm{log}~P$ values have been
    retrieved from the Pubchem database (118 estimated data and 6 substances without data).  Blue line
    indicates the regression model
    ($\mathrm{log~LC}_{50} = 2.86\ensuremath{-0.36}~\mathrm{log}~P$,
    RMSE = 1.47).}
\label{fig:fig3}
\end{figure}
